---
title: "valueBoxes and Sparklines"
description: |
  The value boxes are a good way to get to the point quickly in a dashboard
  showing the relevant value with a big font and eye-catching color. 
  But sometimes it would be better show that relevant number with a 
  related text, sparkline or information to give context of that value, or 
  just to compare it with the previoues ones. 
  In this post we'll show how to add sparklines to valueBox in shinydashboard
  and other related packages.
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 06-29-2020
output:
  distill::distill_article:
    always_allow_html: true
    self_contained: true
    toc: true
draft: false
categories: 
  - shiny
  - highcharts
  - htmlwidgets
editor_options: 
  chunk_output_type: console
preview: images/preview.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source(here::here("_R/blog_setup.R"))

message("this delete the images created in interactive mode (project's root)")
# try(fs::dir_delete("images"))
```

## Introducction

When you create a dashboard to track some information is usual put the important
numbers big and clear at the beginning of the main section. 
For this, {shinydashboard} (https://rstudio.github.io/shinydashboard/) and other dashboard oriented
packages for shiny like {bs4dash}, {argonDash}, {tablerDash} 
- all from https://rinterface.com/ - have an 
implementation to put this type of information know as a _card_ or _value box_.

In some projects like https://jbkunst.shinyapps.io/trd-sttstcs/ I've implemented a
modification of the original `shinydashboard::valueBox`
to complement the information (the BIG number) with a sparkline. Always
that implementation lives as a function in `R/helpers.R` script, so this will
be a good opportunity to share the code, thoughts and ideas. 

```{r, echo=FALSE, layout="l-body", fig.cap="https://twitter.com/jbkunst/status/1246082350922031105"}
tweetrmd::tweet_screenshot("https://twitter.com/jbkunst/status/1246082350922031105")
```

## The basic `valueBox`

Let's start with what we know: The most direct and simple implementation
of a `valueBox` in {shinydashboard}.

```{r shinyapp}
library(shiny)
library(shinydashboard)

vb <-  valueBox(
  value = "1,345",
  subtitle = "Lines of code written",
  icon = icon("code"),
  width = 4,
  color = "red",
  href = NULL)

ui <- dashboardPage(
  dashboardHeader(),
  dashboardSidebar(disable = TRUE),
  dashboardBody(
    fluidRow(
      valueBoxOutput("vbox")
    )
  )
)

server <- function(input, output) {
  
  output$vbox <- renderValueBox(vb)
  
}

# shiny::shinyApp(ui, server, options = list(launch.browser = .rs.invokeShinyPaneViewer))
```

This app look like:

```{r, fig.cap="The most simple value box", echo=FALSE}
# shiny::shinyApp(ui, server, options = list(launch.browser = .rs.invokeShinyPaneViewer))

library(webshot2)

img_file <- "images/app1_vb.png"

appshot(shinyApp(ui, server), file = img_file, zoom = 2, vheight = 500)

knitr::include_graphics(img_file)
```

So here it is,  a simple value box to emphasize information which can have an 
icon/color to give some meaning to the number.

```{r, fig.cap="Apply some zoom", echo=FALSE}
img_file <- "images/vb1.png"

appshot(shinyApp(ui, server), selector = "#vbox", file = img_file, zoom = 2)

knitr::include_graphics(img_file)
```

## Modifying the `valueBox`

Before to modify a value box we need to know what a `valueBox` is, or 
how is it built, so take a look what is inside this funcion.

```{r, eval=FALSE}
valueBox

function(value, subtitle, icon = NULL, color = "aqua", width = 4, href = NULL){
  
  validateColor(color)
  
  if (!is.null(icon))
    tagAssert(icon, type = "i")
  
  boxContent <- div(
    class = paste0("small-box bg-", color),
    div(
      class = "inner",
      h3(value),
      p(subtitle)
      ),
    if (!is.null(icon)) div(class = "icon-large", icon)
    )
  
  if (!is.null(href)) 
    boxContent <- a(href = href, boxContent)
    
  div(
    class = if (!is.null(width)) paste0("col-sm-", width), 
    boxContent
    )
}
```

As we see, the `value` and `subtitle` arguments are the main ones. The other are
just for styling and format. These elements are in `div(class = "inner", ...)`
so here we'll put the new content there.

Let's try to put a title in the top of the div and then use the subtitle as 
an detailed text.


```{r}
valueBox2 <- function(value, title, subtitle, icon = NULL, color = "aqua", width = 4, href = NULL){
  
  shinydashboard:::validateColor(color)
  
  if (!is.null(icon))
    shinydashboard:::tagAssert(icon, type = "i")
  
  boxContent <- div(
    class = paste0("small-box bg-", color),
    div(
      class = "inner",
      tags$small(title),
      h3(value),
      p(subtitle)
      ),
    if (!is.null(icon)) div(class = "icon-large", icon)
    )
  
  if (!is.null(href)) 
    boxContent <- a(href = href, boxContent)
    
  div(
    class = if (!is.null(width)) paste0("col-sm-", width), 
    boxContent
    )
}
```

So the implementation is the same as the original `valueBox` function:

```{r}
vb <- valueBox2(
  value = "1,345",
  title = toupper("Lines of code"),
  subtitle = tagList(HTML("&uarr;"), "25% Since last day"),
  icon = icon("code"),
  width = 4,
  color = "red",
  href = NULL
  )
```


```{r, fig.cap="A modificated value box", echo=FALSE}
# shiny::shinyApp(ui, server, options = list(launch.browser = .rs.invokeShinyPaneViewer))

img_file <- "images/vb2.png"

appshot(shinyApp(ui, server), selector = "#vbox", file = img_file, zoom = 2)

knitr::include_graphics(img_file)
```

The idea of put a title is to help the reader, but we use a small text to 
be _subtle_, usually I like to use a _Small Caps_ font but in this case
I used uppercase for simplicity.

## Adding sparklines

The idea to add a sparkline to the value box is give more information. For example,
if your have a historical data, it would be good put a sparkline to show the previous
behaviour about the data, so you can compare the _BIG_ number with the past.

Let's simulate data to create a sparkline:

```{r}
set.seed(123)

N <- 20

x <- cumsum(rnorm(N)) + 0.5 * cumsum(runif(N))
x <- round(200*x)

df <- data.frame(
  x = sort(as.Date(Sys.time() - lubridate::days(1:N))),
  y = abs(x)
)

df
```

Let's create a sparkline with {highcharter} package:

```{r}
library(highcharter)

hc <- hchart(df, "line", hcaes(x, y), name = "lines of code")  %>% 
  hc_size(height = 100)

hc
```

We have our first sparkline. It's important note the `hc` object will use
the available horizontal (`width`) space, but we need to control the `heigth`.

Now we can create a new `valueBox` function;

```{r}
valueBox3 <- function(value, title, sparkobj = NULL, subtitle, icon = NULL, color = "aqua", width = 4, href = NULL){
  
  shinydashboard:::validateColor(color)
  
  if (!is.null(icon))
    shinydashboard:::tagAssert(icon, type = "i")
  
  boxContent <- div(
    class = paste0("small-box bg-", color),
    div(
      class = "inner",
      tags$small(title),
      h3(value),
      if (!is.null(sparkobj)) sparkobj,
      p(subtitle)
      ),
    if (!is.null(icon)) div(class = "icon-large", icon, style = "z-index; 0")
    )
  
  if (!is.null(href)) 
    boxContent <- a(href = href, boxContent)
    
  div(
    class = if (!is.null(width)) paste0("col-sm-", width), 
    boxContent
    )
}
```

So now:

```{r}
vb <- valueBox3(
  value = "1,345",
  title = toupper("Lines of code"),
  sparkobj = hc,
  subtitle = tagList(HTML("&uarr;"), "25% Since last day"),
  icon = icon("code"),
  width = 4,
  color = "red",
  href = NULL)
```

And the result is:

```{r, fig.cap="A value box with a sparkline using a bad theme", echo=FALSE}
# shiny::shinyApp(ui, server, options = list(launch.browser = .rs.invokeShinyPaneViewer))

img_file <- "images/vb3.png"

appshot(shinyApp(ui, server), selector = "#vbox", file = img_file, zoom = 2, delay = 2)

knitr::include_graphics(img_file)
```

LOL! We need modify the look of the chart. For this we'll create a custom theme.
The general idea is gain space via removing axis and borders. Then have cleaner
look removing grid lines... You know, show _just_ the line.

```{r hc_theme_sparkline_vb}
hc_theme_sparkline_vb <- function(...) {
  
  theme <- list(
    chart = list(
      backgroundColor = NULL,
      margins = c(0, 0, 0, 0),
      spacingTop = 0,
      spacingRight = 0,
      spacingBottom = 0,
      spacingLeft = 0,
      plotBorderWidth = 0,
      borderWidth = 0,
      style = list(overflow = "visible")
    ),
    xAxis = list(
      visible = FALSE, 
      endOnTick = FALSE, 
      startOnTick = FALSE
    ),
    yAxis = list(
      visible = FALSE,
      endOnTick = FALSE, 
      startOnTick = FALSE
    ),
    tooltip = list(
      outside = FALSE,
      shadow = FALSE,
      borderColor = "transparent",
      botderWidth = 0,
      backgroundColor = "transparent",
      style = list(textOutline = "5px white")
    ),
    plotOptions = list(
      series = list(
        marker = list(enabled = FALSE),
        lineWidth = 2,
        shadow = FALSE,
        fillOpacity = 0.25,
        color = "#FFFFFFBF",
        fillColor = list(
          linearGradient = list(x1 = 0, y1 = 1, x2 = 0, y2 = 0),
          stops = list(
            list(0.00, "#FFFFFF00"),
            list(0.50, "#FFFFFF7F"),
            list(1.00, "#FFFFFFFF")
          )
        )
      )
    ),
    credits = list(
      enabled = FALSE,
      text = ""
      )
    )
  
  theme <- structure(theme, class = "hc_theme")
  
  if (length(list(...)) > 0) {
    theme <- hc_theme_merge(
      theme,
      hc_theme(...)
    )
  }
  
  theme
}
```

Testing the theme.

```{r}
hc <- hc %>% 
  hc_add_theme(hc_theme_sparkline_vb()) %>% 
  hc_credits(enabled = FALSE)

hc %>% 
  # emulate the background color of the valueBox
  hc_chart(backgroundColor = "#DD4B39")
```

And trying again:

```{r}
vb <- valueBox3(
  value = "1,345",
  title = toupper("Lines of code"),
  sparkobj = hc,
  subtitle = tagList(HTML("&uarr;"), "25% Since last day"),
  icon = icon("code"),
  width = 4,
  color = "red",
  href = NULL)
```


```{r, fig.cap="A value box with a proper sparkline", echo=FALSE}
# shiny::shinyApp(ui, server, options = list(launch.browser = .rs.invokeShinyPaneViewer))

img_file <- "images/vb4.png"

appshot(shinyApp(ui, server), selector = "#vbox", file = img_file, zoom = 2, delay = 2)

knitr::include_graphics(img_file)
```

## Including an information helper 

Sometimes I like to add helper information icon to explain how to read the 
metric/number if is needed.

This can be done using the icon `"info-cirle"` and using the class `"pull-right"`
to put in the corner.

```{r}
valueBox4 <- function(value, title, sparkobj = NULL, subtitle, info = NULL, icon = NULL, color = "aqua", width = 4, href = NULL){
  
  shinydashboard:::validateColor(color)
  
  if (!is.null(icon))
    shinydashboard:::tagAssert(icon, type = "i")
  
  info_icon <- tags$small(
    tags$i(
      class = "fa fa-info-circle fa-lg",
      title = info,
      `data-toggle` = "tooltip",
      style = "color: rgba(255, 255, 255, 0.75);"
      ),
    class = "pull-right"
    )
  
  boxContent <- div(
    class = paste0("small-box bg-", color),
    div(
      class = "inner",
      tags$small(title),
      if (!is.null(sparkobj)) info_icon,
      h3(value),
      if (!is.null(sparkobj)) sparkobj,
      p(subtitle)
      ),
    if (!is.null(icon)) div(class = "icon-large", icon, style = "z-index; 0")
    )
  
  if (!is.null(href)) 
    boxContent <- a(href = href, boxContent)
    
  div(
    class = if (!is.null(width)) paste0("col-sm-", width), 
    boxContent
    )
}
```

So now:

```{r}
vb <- valueBox4(
  value = "1,345",
  title = toupper("Lines of code"),
  sparkobj = hc,
  subtitle = tagList(HTML("&uarr;"), "25% Since last day"),
  info = "This is the lines of code I've written in the past 20 days! That's a lot, right?",
  icon = icon("code"),
  width = 4,
  color = "red",
  href = NULL
  )
```


```{r, fig.cap="A value box with helper text on hover", echo=FALSE}
# shiny::shinyApp(ui, server, options = list(launch.browser = .rs.invokeShinyPaneViewer))

img_file <- "images/vb5.png"

appshot(shinyApp(ui, server), selector = "#vbox", file = img_file, zoom = 2, delay = 2)

knitr::include_graphics(img_file)
```

_Voilà_ we have a value box with a information helper.

## Demo

Demo section! We can sue other types of charts, texts and colors:

Before the demo, let's ngive a proper name to our function:

```{r valueBoxSpark}
valueBoxSpark <- function(value, title, sparkobj = NULL, subtitle, info = NULL, icon = NULL, color = "aqua", width = 4, href = NULL){
  
  shinydashboard:::validateColor(color)
  
  if (!is.null(icon))
    shinydashboard:::tagAssert(icon, type = "i")
  
  info_icon <- tags$small(
    tags$i(
      class = "fa fa-info-circle fa-lg",
      title = info,
      `data-toggle` = "tooltip",
      style = "color: rgba(255, 255, 255, 0.75);"
      ),
    # bs3 pull-right 
    # bs4 float-right
    class = "pull-right float-right"
    )
  
  boxContent <- div(
    class = paste0("small-box bg-", color),
    div(
      class = "inner",
      tags$small(title),
      if (!is.null(sparkobj)) info_icon,
      h3(value),
      if (!is.null(sparkobj)) sparkobj,
      p(subtitle)
      ),
    # bs3 icon-large
    # bs4 icon
    if (!is.null(icon)) div(class = "icon-large icon", icon, style = "z-index; 0")
    )
  
  if (!is.null(href)) 
    boxContent <- a(href = href, boxContent)
    
  div(
    class = if (!is.null(width)) paste0("col-sm-", width), 
    boxContent
    )
}
```

Some charts:
 
```{r democharts}
hc <- hchart(df, "area", hcaes(x, y), name = "lines of code")  %>% 
  hc_size(height = 100) %>% 
  hc_credits(enabled = FALSE) %>% 
  hc_add_theme(hc_theme_sparkline_vb()) 

hc2 <- hchart(df, "line", hcaes(x, y), name = "Distance")  %>% 
  hc_size(height = 100) %>% 
  hc_credits(enabled = FALSE) %>% 
  hc_add_theme(hc_theme_sparkline_vb()) 

hc3 <- hchart(df, "column", hcaes(x, y), name = "Daily amount")  %>% 
  hc_size(height = 100) %>% 
  hc_credits(enabled = FALSE) %>% 
  hc_add_theme(hc_theme_sparkline_vb()) 
```

Some value boxes:

```{r}
vb <- valueBoxSpark(
  value = "1,345",
  title = toupper("Lines of code written"),
  sparkobj = hc,
  subtitle = tagList(HTML("&uarr;"), "25% Since last day"),
  info = "This is the lines of code I've written in the past 20 days! That's a lot, right?",
  icon = icon("code"),
  width = 4,
  color = "teal",
  href = NULL
  )

vb2 <- valueBoxSpark(
  value = "1,345 KM",
  title = toupper("Distance Traveled"),
  sparkobj = hc2,
  subtitle = tagList(HTML("&uarr;"), "25% Since last month"),
  info = "This is the lines of code I've written in the past 20 days! That's a lot, right?",
  icon = icon("plane"),
  width = 4,
  color = "red",
  href = NULL
  )

vb3 <- valueBoxSpark(
  value = "1,3 Hrs.",
  title = toupper("Thinking time"),
  sparkobj = hc3,
  subtitle = tagList(HTML("&uarr;"), "5% Since last year"),
  info = "This is the lines of code I've written in the past 20 days! That's a lot, right?",
  icon = icon("hourglass-half"),
  width = 4,
  color = "yellow",
  href = NULL
  )
```

Finally the app:


```{r shinyapp2}
library(shiny)
library(shinydashboard)

ui <- dashboardPage(
  dashboardHeader(),
  dashboardSidebar(disable = TRUE),
  dashboardBody(
    fluidRow(
      valueBoxOutput("vbox"),
      valueBoxOutput("vbox2"),
      valueBoxOutput("vbox3")
    )
  )
)

server <- function(input, output) {
  output$vbox <- renderValueBox(vb)
  output$vbox2 <- renderValueBox(vb2)
  output$vbox3 <- renderValueBox(vb3)
}

# shiny::shinyApp(ui, server, options = list(launch.browser = .rs.invokeShinyPaneViewer))
```

This app look like this :)! What do you think?


```{r, fig.cap="Awesome new value boxes", echo=FALSE, layout="l-page"}
# shiny::shinyApp(ui, server, options = list(launch.browser = .rs.invokeShinyPaneViewer))
img_file <- "images/vb6.png"

appshot(shinyApp(ui, server), selector = "#vbox2", file = img_file, zoom = 2, delay = 2)

img_file <- "images/vb7.png"

appshot(shinyApp(ui, server), selector = "#vbox3", file = img_file, zoom = 2, delay = 2)


img_file <- "images/app2_vb.png"

webshot2::appshot(shinyApp(ui, server), file = img_file, zoom = 2, vheight = 500, delay = 4)

knitr::include_graphics(img_file)
```

## Extending the idea

Due the {bs4dash} package is based in the same template of the {shinydashboard}
https://github.com/ColorlibHQ/AdminLTE is direct use the functions. There some
details in the info icon, but you can remove it if you use `icon = NULL`.

```{r}
library(bs4Dash)

ui <- bs4DashPage(
  navbar = bs4DashNavbar(),
  sidebar = bs4DashSidebar(disable = TRUE),
  body = bs4DashBody(
    # this is for use tooltips in the bs4dash package
    tags$script(HTML("setInterval(function(){ $('[title]').tooltip(); }, 1000)")),
    tags$h2("Testing with {bs4dash} package"),
    tags$hr(),
    fluidRow(
      valueBoxOutput("vbox"),
      valueBoxOutput("vbox2"),
      valueBoxOutput("vbox3")
      )
    )
  )

# setTimeout(function(){ $('[title]').tooltip(); }, 3000)
# setInterval(function(){ $('[title]').tooltip(); }, 3000)
# JS("$(function () { $('[title]').tooltip() })")
```


```{r, fig.cap="Work with the {bs4dash} package too", echo=FALSE, layout="l-page"}
# shiny::shinyApp(ui, server, options = list(launch.browser = .rs.invokeShinyPaneViewer))

img_file <- "images/vb8.png"
appshot(shinyApp(ui, server), selector = "#vbox", file = img_file, zoom = 2, delay = 2)


img_file <- "images/vb9.png"
appshot(shinyApp(ui, server), selector = "#vbox2", file = img_file, zoom = 2, delay = 2)

img_file <- "images/vb10.png"
appshot(shinyApp(ui, server), selector = "#vbox3", file = img_file, zoom = 2, delay = 2)

img_file <- "images/app3_vb.png"

webshot2::appshot(shinyApp(ui, server), file = img_file, zoom = 2, vheight = 500, delay = 5)

knitr::include_graphics(img_file)
```

This walkthrough/tutorial will help to implement sparklines in other packages 
as well as other elements such as `infoBox`, or maybe with other htmlwidgets 
like {echarts4r} or {plotly}. So if you make an implementation 
please tell us and share with other people. 
