---
title: "How to: Weather Radials"
description: |
  A short description of the post.
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 03-24-2016
output:
  distill::distill_article:
    self_contained: true
categories: 
  - data-visualization
  - highcharts
editor_options: 
  chunk_output_type: console
preview: images/preview.gif    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(here::here("_R/blog_setup.R"))
```

TLDR: Creating weather radials with [highcharter](http://jkunst.com/highcharter/) and 
[ggplot2](http://docs.ggplot2.org/current/).

I was surfing by the deep seas of the web and I found the *Brice Pierre de la Briere*'s 
[blocks](http://bl.ocks.org/bricedev) and I saw the *weather radials* which originally are a
[poster collection](http://weather-radials.com/). Brice uses D3 and he used D3 very well
and I love D3 but I'm in a rookie level to do something like him. **D3 is not for everybody**
and surely not for me, I would love to lear more but family, work and R has priority over D3 so
how can I do something like that. Well... We have R & highcharter. So let's try. 

We'll use the same data as Brice [https://www.wunderground.com/].

```{r}
library(tidyverse)
library(highcharter)
library(lubridate)

df <- read_csv("http://bl.ocks.org/bricedev/raw/458a01917183d98dff3c/sf.csv")

df

df <- df %>%
  rename_all(str_to_lower) %>% 
  rename_all(str_replace, "\\s+", "_") %>% 
  mutate(
    id = seq(nrow(df)),
    date2 = as.Date(ymd(date)),
    tmstmp = datetime_to_timestamp(date2),
    month = month(ymd(date))
  )

dsmax <- df %>%
  select(tmstmp, max_temperaturec)
 
dsmin <- df %>% 
  select(tmstmp, min_temperaturec)
```


## First try

Here we test and chart the data in the most simple way. A line time.
 
```{r}
hc <- highchart() %>% 
  hc_chart(
    type = "line"
    ) %>%
  hc_xAxis(
    type = "datetime",
    tickInterval = 30 * 24 * 3600 * 1000,
    labels = list(format = "{value: %b}")
  ) %>% 
  hc_yAxis(
    min = 0,
    labels = list(format = "{value} C")
  ) %>% 
  hc_add_series(dsmax, type = "line", hcaes(tmstmp, max_temperaturec), name = "max") %>% 
  hc_add_series(dsmin, type = "line",  hcaes(tmstmp, min_temperaturec), name = "min")

hc
```

Everything seems fine. We now a little bit about the data 

## To polar coordinates and columnrange!

Now, we need to change to polar cordinates to get the "radial" part, this is
get using the `polar = TRUE`  argument in the `hc_chart` function. Then add
the data in the `columnrange` type.
  
```{r}
# Some tooltips to make it a little *intercative*
x <- c("Min", "Mean", "Max")
y <- sprintf("{point.%s}", tolower(x))
tltip <- tooltip_table(x, y)

hc <- hchart(
  df,
  type = "columnrange",
  hcaes(x = date2, low = min_temperaturec, high = max_temperaturec, color = mean_temperaturec),
  showInLegend = FALSE
  ) %>% 
  hc_chart(
    polar = TRUE
  ) %>%  
  hc_xAxis(
    gridLineWidth = 0.5,
    type = "datetime",
    tickInterval = 30 * 24 * 3600 * 1000,
    labels = list(format = "{value: %b}")
  ) %>% 
  hc_yAxis(
    max = 30,
    min = -10,
    labels = list(format = "{value} C"),
    showFirstLabel = FALSE
    ) %>% 
  hc_tooltip(
    useHTML = TRUE,
    headerFormat = as.character(tags$small("{point.x:%d %B, %Y}")),
    pointFormat = tltip
  )

hc
```

Yay :D! A beautiful chart same as the d3 version and only using R. So sweet!
I'm happy with the result. This is not a standar chart but is
a king of *artistic*. What do you think? Any other examples to test
this type of chart?

## Bonus Track: ggplot2 version

It's *really really* easy to do this type of chart in ggplot2 using
`geom_linerange` and `geom_polar`:

```{r}
library(scales)

ggplot(
  df,
  aes(
    date2,
    ymin = min_temperaturec,
    ymax = max_temperaturec,
    color = mean_temperaturec
  )
) +
  geom_linerange(size = 1.3, alpha = 0.75) +
  scale_color_viridis_c(NULL, option = "A") +
  scale_x_date(labels = date_format("%b"), breaks = date_breaks("month")) +
  ylim(-10, 35) +
  labs(
    title = "San Francisco Wather Radial",
    subtitle = "It would be nice if someone do this with the animation package",
    caption = "Other example for ggplot2 vs base #boring but #fun",
    x = NULL,
    y = NULL
  ) +
  coord_polar() +
  theme(legend.position = "bottom")
```

Nice!

Searching I found someone do 
[this](https://www.quora.com/R-programming-language/What-is-the-most-elegant-plot-you-have-made-using-ggplot2):

> Always exist someone who did what you did before you.

At least I share the code! :D.

```{r, echo=FALSE}
# knitr::include_graphics("http://giphy.com/gifs/how-i-met-your-mother-barney-stinson-high-five-13IGxUKu7ucp68")
```
