---
title: "rchess a Chess Package for R"
description: |
  Finally R have a chess package! Let's see what
  can it do.
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 10-07-2015
output:
  distill::distill_article:
    self_contained: true
editor_options: 
  chunk_output_type: console
categories: 
  - package
  - htmlwidgets
preview: images/preview.png  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A long time ago I wonder for some R package for chess. Python, JS, Java, C have 
chess libraries, why R not? But then the [htmlwidgets](https://github.com/ramnathv/htmlwidgets) 
and [V8](https://github.com/jeroenooms/v8) packages were born and now possibilities 
are *almost* endless. Anyone can make a wrapper for a JS library :D. And this is
how [rchess](https://github.com/jbkunst/rchess) came out.

The [rchess](https://github.com/jbkunst/rchess) package is a chess move, 
generation, validator; piece placement, movement, and check/checkmate/stalemate
detection. All this powered by [V8](https://github.com/jeroenooms/v8) package 
and [chessjs](https://github.com/jhlywa/chess.js) javascript library.

The main parts about this package are:

- [V8](https://github.com/jeroenooms/v8) package and [chessjs](https://github.com/jhlywa/chess.js) javascript library. So this
is not about performance but have the things done (just like Just [@hrbrmstr](http://datadrivensecurity.info/blog/posts/2015/Oct/getting-into-the-zones-with-r-jsonlite/)
saids).
- [R6](https://github.com/wch/R6/) package for the OO system.
- [htmlwidgets](https://github.com/ramnathv/htmlwidgets) package and [chessboardjs](http://chessboardjs.com/) javascript library.

Now let's take a look what this package can do.

```{r}
# install.packages("rchess")
# devtools::install_github("jbkunst/rchess")
library(rchess)
chss <- Chess$new()

plot(chss)
```

WHOA chessboardjs have a nice look. Right?

The basic usage is the same as the [chessjs](https://github.com/jhlywa/chess.js) api. I'm just implemented only the calls to
this api. Anyway, to see the possibles moves you can use the *moves* method and if you make a move use the *move*
function.

```{r}
chss$moves()

chss$moves(verbose = TRUE)

chss$move("a3")
```

We can also concate  moves, see who plays in the next turn, see the history, etc.:

```{r}
chss$move("e5")$move("f4")$move("Qe7")$move("fxe5")

plot(chss)
```


```{r}
chss$turn()

chss$history(verbose = TRUE)

chss$history()
```

A must have funcionality is load games from PGN and FEN notations:

```{r}
chssfen <- Chess$new()
 
fen <- "rnbqkbnr/pp1ppppp/8/2p5/4P3/8/PPPP1PPP/RNBQKBNR w KQkq c6 0 2"
 
chssfen$load(fen)
```

(Shh you can also plot the chesboard with ggplot. I know I need change that unicode style pieces ;) ). 

```{r}
plot(chssfen, type = "ggplot")
```

Now load PGN string:

```{r}
pgn <- system.file("extdata/pgn/kasparov_vs_topalov.pgn", package = "rchess")
pgn <- readLines(pgn, warn = FALSE)
pgn <- paste(pgn, collapse = "\n")

chsspgn <- Chess$new()

chsspgn$load_pgn(pgn)

cat(chsspgn$pgn())

head(chsspgn$history(verbose = TRUE), 20)

plot(chsspgn)
```

And that's it. If you want to check more funcionalities check the
[github repository](https://github.com/jbkunst/rchess) or 
[this](http://rpubs.com/jbkunst/rchess2) document.

Bye mate!
