---
title: "Minimalistic toolptips with highcharter/highcharts"
description: |
  One of the many post to remember things that usually forget. Yes, this blog
  is for that.
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 01-07-2021
output:
  distill::distill_article:
    self_contained: true
categories: 
  - minipost
  - highcharts
  - highcharter      
editor_options: 
  chunk_output_type: console
preview: images/preview.jpeg
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Time ago we make some _really nice but complex_ tooltips showing charts from
the selected points (in the chart) using `{highcharter}` package -and highcharts-
in the [using tooltips in unexpected ways](/..2019-02-04-using-tooltips-in-unexpected-ways/) 
post. But sometimes we need less: a tiny toolip, maybe for the available space, or requirements.

So in this post we'll detail the parameters in `highcharter::hc_tooltip` function
used in the [valuebox and sparklines](../2020-06-26-valuebox-and-sparklines) post 
to get what we want.

First, let's try a normal chart with a traditional tooltip.

```{r}
library(highcharter)
library(tidyverse)

hc <- hchart(citytemp_long, "line", hcaes(month, temp, group = city)) %>% 
  hc_add_theme(hc_theme_538()) 

hc
```

Did you see the tooltip? Is just a normal one, nothing special, it works :). Now, 
what we came for, but before change arguments let's make clear what we'll change:

- The main one is give the text outline effect to the text.
- Change the `pointFormat` argument removing the ":" in the text because
the `textOutline` effect don't work very well with this character.
- Change the tooltip background setting the color to transparent.
- Remove the border setting the `borderWidth` to 0.
- Remove the shadow.

And voila.

```{r}
hc %>% 
  hc_tooltip(
    pointFormat = "<span  style='color: {series.color}'> {series.name} <b>{point.y}</b><br/></span>",
    shadow = FALSE,
    backgroundColor = "transparent",
    style = list(textOutline = "3px #404040"),
    borderColor = "transparent",
    borderWidth = 0
  )
```



