---
title: "twitter"
description: |
  A short description of the post.
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 02-05-2021
output:
  distill::distill_article:
    self_contained: true
draft: true
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.



```{r}
library(tidyverse)
library(rtweet)
library(lubridate)
library(highcharter)
```


```{r, echo=FALSE}
# I know this is unsercure but hey, what's the worst that can happen?
token <- create_token(
  app = "jbktests",
  consumer_key = "JXG4ZCzx4oDpbSGh6bU1OwZFa",
  consumer_secret = "JpOOkA44d9EHd2gvznG0yxSywR4yzQzbsxKlIDe2soS8Kd8KwU",
  access_token = "7220752-P9Qvy2u2pqY2NAsmVs3j8AebkuL0zuN2zvBlc5BAcz",
  access_secret = "zhzaTJqQBYGORWEcysi3oojbp4WkYH9TupJrNPUBQ4S3l"
)
```
For download tweets I followed the instructions from https://cran.r-project.org/web/packages/rtweet/vignettes/auth.html

```{r, eval=FALSE}
token <- create_token(
  app = "jbktests",
  consumer_key = "consumerkey",
  consumer_secret = "consumer_secret",
  access_token = "accesstokena98ds087",
  access_secret = "accesssecreta98ds087"
)
```

After get the tweets we filter the data and make some transformations. Then
we'll get the top tweets in terms of the favorite count to make the annotations
in our chart.


```{r}
# in interactive mode you can avoid/remove the token argument
tweets_rstats_path <- here::here("data/rstats-tweets/tweets_rstats.rds")

if(!file.exists(tweets_rstats_path)) {
  
  tweets_rstats <- rtweet::search_tweets(
    q = "#rstats -#javascript -#python -#bigdata -#rbloggers -#datascience",
    n = 18000, 
    retryonratelimit = TRUE,
    include_rts = FALSE
    )
  
  saveRDS(tweets_rstats, tweets_rstats_path)
  
} else {
  
   tweets_rstats <- readRDS(tweets_rstats_path)
   
}

tweets_rstats
```


Order the user using favorite_count

```{r}
users <- tweets_rstats %>% 
  group_by(user = screen_name, user_id) %>% 
  summarise(
    favorite_count = sum(favorite_count),
    retweet_count = sum(retweet_count),
    .groups = "drop"
  ) %>% 
  arrange(desc(favorite_count), desc(retweet_count)) %>% 
  mutate(rank = row_number()) %>% 
  select(rank, everything())

users
```


```{r}
progress_sleep <- function(time, ...) {
  
  time <- round(time) + 1
  
  pb <-  progress::progress_bar$new(total = time, ...)
  # format = "  waiting [:bar] :percent eta: :eta"
  
  for (i in 1:time) {
    pb$tick()
    Sys.sleep(1)
  }
  
}

get_friends_rl <- function(name = "CNN"){
  
  message(name)
  
  rl <- rtweet::rate_limit(query = "get_friends", token = token)
  
  if(rl$remaining == 0) {
    
    stringr::str_glue(
      "Waiting until { reset_at } (~{mins} mins)",
      reset_at = rl$reset_at,
      mins     = round(as.numeric(rl$reset, units = "mins") + 5/60, 2)
      ) %>% 
      message()
    
    as.numeric(rl$reset, units = "secs") %>% 
      ceiling() %>% 
      magrittr::add(5) %>% 
      progress_sleep(format = "[:bar] :percent eta: :eta")
    }
  
  rtweet::get_friends(users = name, retryonratelimit = TRUE, token = token)

}

get_friends_rl("jbkunst")
get_friends_rl("juliasilge")

```


```{r}
TOP <- nrow(users)

users_friends_path <- here::here("data/rstats-tweets/tweets_users_friends.rds")

if(!file.exists(users_friends_path)) {
  
  users_friends <- users %>% 
    slice_head(n = TOP) %>% 
    pull(user) %>% 
    map_df(get_friends_rl)

  saveRDS(users_friends, users_friends_path)
  
} else {
  
   users_friends <- readRDS(users_friends_path)
   
}

users_friends
users_friends %>% count(user)
users_friends %>% count(user_id)

users
```


```{r}
users2 <- users_friends %>% 
  count(user_id, sort = TRUE) 

users2

users2 <- users2 %>% 
  anti_join(users %>% slice_head(n = TOP), by = c("user_id")) %>% 
  slice_head(n = TOP) 

users2 <- users2 %>% 
  pull(user_id) %>% 
  rtweet::lookup_users() %>% 
  select(user_id, user = screen_name) %>% 
  left_join(users2, by = "user_id")



inner_join(users, users2)

bind_rows(users, users2) %>% 
  filter(str_detect(user, "had"))
```


```{r}
users_friends2_path <- here::here("data/rstats-tweets/tweets_users_friends2.rds")

if(!file.exists(users_friends2_path)) {
  
  users_friends2 <- users2 %>% 
    slice_head(n = TOP) %>% 
    pull(user) %>% 
    map_df(get_friends_rl)

  saveRDS(users_friends2, users_friends2_path)
  
} else {
  
   users_friends2 <- readRDS(users_friends2_path)
   
}

users_friends2
```

```{r}
users_friends  %>% count(user)
users_friends2 %>% count(user)

users_friends  %>% count(user_id)
users_friends2 %>% count(user_id)

bind_rows(users_friends, users_friends2) %>% 
  distinct(user) %>% 
  filter(str_detect(user, ""))

```



