---
title: The Adevnture on Shiny Server, Rasperry Pi and possibly Arduino (part 1/n)
description: |
  .
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: '2021-07-21'
output:
  distill::distill_article:
    always_allow_html: true
    self_contained: true
    toc: true
draft: true
categories: 
  - shiny
  - shiny-server
  - rasperry-pi
  - dont-forget
editor_options: 
  chunk_output_type: console
preview: images/preview.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source(here::here("_R/blog_setup.R"))

message("this delete the images created in interactive mode (project's root)")
# try(fs::dir_delete("images"))
```

```{r, echo=FALSE}
invisible(Sys.setlocale("LC_TIME", "English"))
```

## Why? Tell me why!

Time ago I see the post from Antonio Vazquez Brust [Code live from RStudio, and share it with the World in real time](https://bitsandbricks.github.io/post/compartiendo-c%C3%B3digo-en-vivo-con-el-mundo-desde-rstudio/)
who uses the {livecode} package and the https://ngrok.com/ service to _expose_
a script a file to the internet.

So I think: instead of share a script, you/we can share an Shiny app `r emo::ji("smile")`,
using a RasperryPi as server and to make it funnier we can track data and show 
historic and real time data from sensors using arduino.

Why a Rasperry-Pi? I've used https://www.digitalocean.com/ and https://www.shinyapps.io/
to serve shiny apps and have relative similar to Rasperry-Pi in terms of hardware. Obviously all these options
have pros and cons, and not are the same service but I think is a good chance
to explore options and get some comparision:

```{r, echo=FALSE, layout='l-page', fig.cap="Prices at Julyu 2021."}
tibble::tibble(
  `Category`   = c("RAM", "CPU", "Memory", "Price", "Purpose", "Deploy"),
  `Digital Ocean (basic)` = c("1GB", "Shared", "25GB", "72 USD/year", "General", "Via terminal"),
  `shinyapps.io (starter)` = c("", "", "", "100 USD/year", "Only Shinyapps", "One click from rstudio"),
  `Rasperry PI` = c("8GB", "1.4 Quadcore", "32GB", "From 35USD to 100USD (w/ accesories) one time", "General", "Via terminal")
) %>% 
  knitr::kable() %>% 
  kableExtra::column_spec(column = 1, bold = TRUE) 
```


