---
title: "Favorite and retweets counts from @year_progress account"
description: |
  I had this doubt a long time ago, and the results make sense with the expected... or maybe not?!
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 01-06-2021
output:
  distill::distill_article:
    self_contained: true
editor_options: 
  chunk_output_type: console
preview: images/preview.png
---


```{r setup, include=FALSE}
source(here::here("_R/blog_setup.R"))
```

You know, is the end of the year and more than one of your contacts retweet:

```{r echo=FALSE}
tweetrmd::tweet_screenshot("https://twitter.com/dsquintana/status/1343481829819633665")
```

Surely you remembered that in the middle of the year you saw the 50% retweet. 
So why not see the number of favorites throughout the year? To do this we need
some packages! The main one in this post will be `{rtweet}`, because..,
we need tweets.

```{r}
library(tidyverse)
library(rtweet)
library(lubridate)
library(highcharter)
```


```{r, echo=FALSE}
# I know this is unsercure but hey, what's the worst that can happen?
token <- create_token(
  app = "jbktests",
  consumer_key = "JXG4ZCzx4oDpbSGh6bU1OwZFa",
  consumer_secret = "JpOOkA44d9EHd2gvznG0yxSywR4yzQzbsxKlIDe2soS8Kd8KwU",
  access_token = "7220752-P9Qvy2u2pqY2NAsmVs3j8AebkuL0zuN2zvBlc5BAcz",
  access_secret = "zhzaTJqQBYGORWEcysi3oojbp4WkYH9TupJrNPUBQ4S3l"
)
```
For download tweets I followed the instructions from https://cran.r-project.org/web/packages/rtweet/vignettes/auth.html

```{r, eval=FALSE}
token <- create_token(
  app = "jbktests",
  consumer_key = "consumerkey",
  consumer_secret = "consumer_secret",
  access_token = "access_token",
  access_secret = "access_secret"
)
```

After get the tweets we filter the data and make some transformations. Then
we'll get the top tweets in terms of the favorite count to make the annotations
in our chart.


```{r}
# in interactive mode you can avoid/remove the token argument
dtweets <- rtweet::get_timeline("year_progress", n = 400, token = token)

dtweets <- dtweets %>% 
  filter(between(year(created_at), 2020, 2020)) %>% 
  mutate(created_at = as.Date(created_at)) 

dtweets <- dtweets %>% 
  distinct(created_at, .keep_all = TRUE)

dtops <- dtweets %>% 
  top_n(5, favorite_count) %>% 
  select(created_at, text, favorite_count, retweet_count)

dtops
```


```{r}
dtext <- dtweets %>% 
  select(created_at, text) %>% 
  mutate(text = str_extract(text, "[0-9]+.*"))

data <- dtweets %>% 
  select(created_at, Favorites = favorite_count, Retweets = retweet_count) %>% 
  gather(type, count, -created_at) %>% 
  arrange(created_at) %>% 
  left_join(dtext, by = "created_at")
```

## Fisrt chart

```{r}
hchart(data, "line", hcaes(created_at, count, group = type)) 
```

<aside>
"keikaku" means "plan" in Japanese
</aside>

Ok, all according to keikau.

Here we can see some peaks, and yes, the peaks are multiples of 5: 10, 15, 20, 75, etc.
and some relevant points to, the top ones: 0, 50, 69, 98, 99.

## Annotations

It would be useful to add annotations to our graph with the relevant points.

```{r}
df_to_annotations_labels <- function(df, xAxis = 0, yAxis = 0) {
  
  stopifnot(hasName(df, "x"))
  stopifnot(hasName(df, "y"))
  stopifnot(hasName(df, "text"))
  
  df %>% 
    rowwise() %>% 
    mutate(point = list(list(x = x, y = y, xAxis = 0, yAxis = 0))) %>% 
    select(-x, -y) %>% 
    list_parse()
  
}


top_annotations <- dtops %>%
  mutate(created_at = datetime_to_timestamp(created_at)) %>% 
  select(x = created_at, y = favorite_count, text) %>% 
  df_to_annotations_labels()
```

## Final chart

To our final chart we'll add some details as a better tooltip, title and the
usual obvious and importan stuff.

```{r, layout="l-page"}
hchart(data, "line", hcaes(created_at, count, group = type)) %>% 
  hc_title(text = "Favorite and retweets counts from @year_progress account in 2020") %>% 
  hc_tooltip(
    shared = TRUE, 
    headerFormat = '{point.key}<br/>Progress: {point.text}<br/><table>',
    table = TRUE
  ) %>% 
   hc_annotations(
    list(
      labelOptions = list(
        shape = "connector",
        align = "right",
        justify = FALSE,
        crop = TRUE,
        style = list(fontSize = "0.8em", textOutline = "1px white")
        ),
      labels = top_annotations
      )
    ) %>% 
  hc_yAxis(title = list(text = "Count")) %>% 
  # from http://jsfiddle.net/kka8eyg5/3/
  hc_xAxis(
    title = list(text = NA),
    endOnTick = FALSE,
    labels = list(
      staggerLines = 1,
      formatter = JS("function () { return Highcharts.dateFormat('%B', this.value); }"),
      tickPositioner = JS("function () {
                var positions = [],
                    tick = Math.floor(this.dataMin),
                    increment = 1000 * 3600 * 24 * 91.5; // 3 months

                for (tick; tick <= this.dataMax; tick += increment) {
                    positions.push(tick);
                }
                if (positions.indexOf(this.dataMax) == -1) positions.push(this.dataMax);
                return positions;
            }")
    )
  )
  
```

What do you think? Was the result expected?
