---
title: "Visualizing sort algorithms with ggplot2"
description: |
  Just learning ggplot2 with inspiration from looking at Sorting Algorithms by Mike Bostock
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 09-25-2015
output:
  distill::distill_article:
    self_contained: true
categories:
  - ggplot2
  - data-visualization
editor_options: 
  chunk_output_type: console
preview: images/preview.png    
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(here::here("_R/blog_setup.R"))
```

Have you read [Visualizing Algorithms](http://bost.ocks.org/mike/algorithms/) 
by Mike Bostock? It's a *pure gold post*. In that post Mike show a *static* 
representation of a sort algorithm and obvious it will fun to replicate that image
with ggplot2 so here we go.

We need some sorts algorithms. In [this](http://faculty.cs.niu.edu/~hutchins/csci230/sorting.htm) link you can
see some algorithms. 
 
We start with Insertion sort:

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)

theme_set(theme_void())

insertion_sort_steps <- function(x  = sample(1:15)){
  
  msteps <- matrix(data = x, ncol = length(x))
  
  for (i in 2:length(x)) {
    
    j <- i
    
    while ((j > 1) && (x[j] < x[j - 1])) {
      
      temp <- x[j]
      x[j] <- x[j - 1]
      x[j - 1] <- temp
      j <- j - 1
      
      msteps <- rbind(msteps, as.vector(x))
      
    }
  }
  
  msteps
  
}
```

Now to test it and see what the function do:

```{r}
set.seed(12345)

x <- sample(seq(4))

x

msteps <- insertion_sort_steps(x)

as.data.frame(msteps)
```

Every *row* is a step in sort the algorithm (a partial sort). This matrix is a hard to plot so 
we need a nicer structure. We can transform the matrix to a *data_frame* 
with the information of every *position* of every *element* in each *step*. 

```{r}
sort_matix_to_df <- function(msteps){
  
  df <- as.data.frame(msteps, row.names = NULL)
  
  names(df) <- seq(ncol(msteps))
  
  df_steps <- df %>%
    tbl_df() %>% 
    mutate(step = seq(nrow(.))) %>% 
    gather(position, element, -step) %>%
    arrange(step)
  
  df_steps
  
}
```

And we apply this function to the previous *steps matrix*.

```{r}
df_steps <- sort_matix_to_df(msteps)

head(df_steps, 10)
```

The next step will be plot the data frame.

```{r}
plot_sort <- function(df_steps, size = 5, color.low = "#D1F0E1", color.high = "#524BB4"){
  
  ggplot(df_steps,
         aes(step, position, group = element, color = element, label = element)) +  
    geom_path(size = size, alpha = 1, lineend = "round") +
    scale_colour_gradient(low = color.low, high = color.high) +
    coord_flip() + 
    scale_x_reverse() + 
    theme(legend.position = "none")
  
}
```

Now compare this:

```{r}
as.data.frame(msteps)
```

With:

```{r}
plot_sort(df_steps, size = 6) + 
  geom_text(color = "white", size = 4)
```

It works, so we can now scroll! 

```{r fig.height=30}
sample(seq(50)) %>% 
  insertion_sort_steps() %>% 
  sort_matix_to_df() %>% 
  plot_sort(size = 2.0)
```

Now try with other sort algorithms:

Bubble sort:

```{r}
bubble_sort_steps <- function(x = sample(1:15)){
  
  msteps <- matrix(data = x, ncol = length(x))
  
  for (i in 1:(length(x) - 1)) {
    
    for (j in 1:(length(x) - 1)) {
      
      if (x[j] > x[j + 1]) {
        temp <- x[j]
        x[j] <- x[j + 1]
        x[j + 1] <- temp
      }
      
      msteps <- rbind(msteps, as.vector(x))
      
    }
  }
  
  msteps
  
}
```

Selection sort:

```{r}
selection_sort_steps <- function(x = sample(1:15)){
  
  msteps <- matrix(data = x, ncol = length(x))
  
  for (i in 1:(length(x) - 1)) {
    
    smallsub <- i
    
    for (j in (i + 1):(length(x) - 0)) { # Is not '- 1' like website
      
      if (x[j] < x[smallsub]) {
        smallsub <- j
      }
    }
    
    temp <- x[i]
    x[i] <- x[smallsub]
    x[smallsub] <- temp
    
    msteps <- rbind(msteps, as.vector(x))
    
  }
  
  msteps
  
}

```

Now test with a longer vector:

```{r}
n <- 50
x <- sample(seq(n))

big_df <- rbind(
  x %>% selection_sort_steps() %>% sort_matix_to_df() %>% mutate(sort = "Selection Sort"),  
  x %>% insertion_sort_steps() %>% sort_matix_to_df() %>% mutate(sort = "Insertion Sort"),
  x %>% bubble_sort_steps() %>% sort_matix_to_df() %>% mutate(sort = "Bubble Sort")
)

head(big_df)

big_df %>%
  group_by(sort) %>% 
  summarise(steps = n())
```


```{r  fig.width=12, fig.height=10}
ggplot(
  big_df,
  aes(step, position, group = element, color = element, label = element)
  ) +  
  geom_path(size = 0.8, alpha = 1, lineend = "round") +
  scale_colour_gradient(low = "#c21500", high = "#ffc500") + # http://uigradients.com/#Kyoto
  facet_wrap(~sort, scales = "free_x", ncol = 1) +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "transparent", linetype = 0),
    strip.text = element_text(size = 8)
    )

```

Or we can plot vertically using the viridis palette from the [viridis package](https://github.com/sjmgarnier/viridis) : 

```{r  fig.width=12, fig.height=10}
ggplot(
  big_df,
  aes(position, step, group = element, color = element, label = element)
  ) +  
  geom_path(size = 1, alpha = 1, lineend = "round") +
  scale_colour_gradientn(colours = viridis_pal()(n)) +
  facet_wrap(~sort, scales = "free_y", nrow = 1) +
  scale_y_reverse() +
  theme(
    legend.position = "none",
    strip.background = element_rect(fill = "transparent", linetype = 0),
    strip.text = element_text(size = 8)
    )
```

And that's it. If you write/implement another sort algorithm in this way let me know to view it ;). 

References:

1. http://bost.ocks.org/mike/algorithms/
1. http://faculty.cs.niu.edu/~hutchins/csci230/sorting.htm
1. http://corte.si/posts/code/visualisingsorting/
1. http://uigradients.com/#Kyoto
1. http://algs4.cs.princeton.edu/21elementary/

```{r featured_image, echo=FALSE, fig.width=5, fig.height=3, fig.show="hide"}
df <- data.frame(x = seq(1:20))
ggplot(df) + 
  geom_bar(
    aes(x = x, y = x, fill = factor(x)),
    stat = "identity" ,width = 0.5
    ) +
  scale_fill_manual(values = viridis_pal(option = "B")(20)) +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "transparent", colour = "transparent"),
    strip.text = element_text(size = 7)
    )

ggsave(filename = "images/preview.png")
```