---
title: "Using tooltips in unexpected ways"
description: |
  Highcharts tooltips are one of most underated features. In the tooltips you
  can easily render html, that means we can put images, tables even charts. 
  So the possibilities are almost limitless.
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 02-04-2019
output:
  distill::distill_article:
    self_contained: true
categories: 
  - data-visualization
  - highcharts    
editor_options: 
  chunk_output_type: console
preview: images/preview.png  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(highcharter)
options(highcharter.theme = hc_theme_smpl())
```

In highcharter R package there is tooltip_chart helper function to put charts 
in the tooltip. Sadly te function is not that easy to use but is not 
impossible :).

The example in the documentation is:

```{r, layout="l-page"}
library(tidyverse)
library(gapminder)
data(gapminder, package = "gapminder")

gp <- gapminder %>%
  arrange(desc(year)) %>%
  distinct(country, .keep_all = TRUE)

gp2 <- gapminder %>%
  select(country, year, pop) %>% 
  nest(-country) %>%
  mutate(
    data = map(data, mutate_mapping, hcaes(x = year, y = pop), drop = TRUE),
    data = map(data, list_parse)
    ) %>%
  rename(ttdata = data)

gptot <- left_join(gp, gp2, by = "country")

hchart(
  gptot,
  "point",
  hcaes(lifeExp, gdpPercap, name = country, size = pop, group = continent)
  ) %>%
  hc_yAxis(type = "logarithmic") %>% 
  # here is the magic (inside the function)
  hc_tooltip(useHTML = TRUE, pointFormatter = tooltip_chart(accesor = "ttdata"))
```

Now we'll use a donut chart and try to place the tooltip inside to give it 
importance:

```{r, layout="l-page"}
donutdata <- gp %>% 
  group_by(continent) %>% 
  summarise(pop = sum(pop/1e6)*1e6)

hchart(donutdata, "pie", hcaes(name = continent, y = pop), innerSize = 300)
```

> Just according to keikaku.

The donut is ready. So now we need a detailed data from each continent to
show in the tooltip.

```{r, layout="l-page"}
donutdata2 <- gp %>% 
  select(continent, lifeExp, gdpPercap) %>% 
  nest(-continent) %>% 
  mutate(
    data = map(data, mutate_mapping, hcaes(x = lifeExp, y = gdpPercap), drop = TRUE),
    data = map(data, list_parse)
    ) %>%
  rename(ttdata = data) %>% 
  left_join(donutdata)

hc <- hchart(
  donutdata2,
  "pie",
  hcaes(name = continent, y = pop),
  innerSize = 375
  )

hc %>% 
  hc_tooltip(
    useHTML = TRUE,
    headerFormat = "<b>{point.key}</b>",
    pointFormatter = tooltip_chart(
      accesor = "ttdata",
      hc_opts = list(
        chart = list(type = "scatter"),
        plotOptions = list(scatter = list(marker = list(radius = 2)))
        ),
      height = 225
      ),
    positioner = JS(
      "function () {
      
        /* one of the most important parts! */
        xp =  this.chart.chartWidth/2 - this.label.width/2
        yp =  this.chart.chartHeight/2 - this.label.height/2
      
        return { x: xp, y: yp };
      
      }"),
    shadow = FALSE,
    borderWidth = 0,
    backgroundColor = "transparent",
    hideDelay = 1000
    )
```

What do you think? Maybe it's a kind of overkill, but hey, it's up to you 
to (ab)use it or not! Another more subtle alternative can be put text, i.e,
the normal tooltip but with a bigger size, like a knob chart.








