---
title: "Simulating and watching autoregressive processes (no {shiny} required)"
description: |
  Always love time series when I studied statistics. This is a post to honor the time series.
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 06-18-2021
output:
  distill::distill_article:
    self_contained: true
    toc: true
    toc_float: true
categories:
  - minipost
  - stats
  - highcharts
  - highcharter
  - time-series
  - data-visualization
editor_options: 
  chunk_output_type: console
preview: images/preview.gif
---

```{r setup, include=FALSE}
source(here::here("_R/blog_setup.R"))
```

# The model

An autoregressive model of order $p$, $AR(p)$, is defined as:

$${\displaystyle X_{t}=c+\sum _{i=1}^{p}\varphi _{i}X_{t-i}+\varepsilon _{t}\,}$$

Where $\varepsilon_t$ is a _White Noise_, $c$, $\varphi_i$ parameters of the model. 
The structure of this model is easy of understand: The next value $X_t$ is a linear combination of the past values plus a random noise -or innovation-.

In R we can simulate this model with `stats::arima.sim` function for example,
if we want simulate an $AR(1)$ with $\varphi _{1} = \varphi = 0.9$  with $0$ mean:

$${\displaystyle X_{t}= \varphi X_{t-i} + \varepsilon _{t} = 0.9 \times X_{t-i} + \varepsilon _{t}\,}$$
We code:

```{r}
ar_model <-  list(ar = c(0.9))

arima.sim(model = ar_model, n = 20)
```

# Interactivity

There is a popular example in highcharts: _Dynamic Update_ where each second
a random value is added to the line series https://www.highcharts.com/demo/dynamic-update.

The key part of this example is the next function which every second generate a 
random value and then is added to the series.

```{js, eval=FALSE}
// This is Javascript code :) :/ :S XD
function () {
    // set up the updating of the chart each second
    var series = this.series[0];
    setInterval(function () {
        var x = (new Date()).getTime(), // current time
            y = Math.random();
        series.addPoint([x, y], true, true);
    }, 1000);
}
```

We'll take the example as a template and change to store the last values
of the process and generate the random noise and the new value. To
get the linear combination between the past values and the autoregressive 
parameter we used the dot product. For the generation of a normal random
number we can use the the Box-Muller transform.

<aside>
_Dot product of two arrays in Javascript_
https://stackoverflow.com/a/64816824/829971

_Generating random normal number in Javascript_
https://stackoverflow.com/a/36481059/829971

</aside>

Let's create some

```{r}
library(highcharter)
library(stringr)

# model
ar <-  c(0.85, -0.1, 0.2)

# time to update the chart in seconds
time <- 1

# the first values of the model
ts <- as.vector(arima.sim(model = list(ar = ar), n = 10))
ts <- round(ts, 3)
ts
```

The function would be something like this:


```{r}
load_fn <- "function () {{

	dot = (a, b) => a.map((x, i) => a[i] * b[i]).reduce((m, n) => m + n);
  function randn_bm() {{
  	var u = 0, v = 0;
    while(u === 0) u = Math.random(); //Converting [0,1) to (0,1)
    while(v === 0) v = Math.random();
    return Math.sqrt( -2.0 * Math.log( u ) ) * Math.cos( 2.0 * Math.PI * v );
   }}
  
	var dat = [{ data }];
  var ar = [{ ar }];
  dat = dat.slice(-ar.length);
  
  // set up the updating of the chart each second
  var series = this.series[0];
  
  setInterval(function () {{
      
      console.log(dat);
      var new_value = dot(dat, ar) + randn_bm();
      new_value = Math.round(1000 * new_value)/1000
      console.log(new_value)
      dat.shift(); 
      dat.push(new_value);
      series.addPoint([new_value]);  
      
      //if (series.data.length < 500) {{
      //  series.addPoint([new_value], true, false);
      //}} else {{
      //  series.addPoint([new_value], true, true);
      //}}
      
  }}, { time });
  
}}"

load_fn_glued <- str_glue(
  load_fn,
  data = str_c(ts, collapse = ","),
  ar = str_c(ar, collapse = ","),
  time = time * 1000
)

load_fn_glued
```


```{r}
highchart() %>% 
  hc_add_series(data = ts) %>% 
  hc_chart(
    events = list(load = JS(load_fn_glued)), 
    animation = list(duration = time*1000/2)
    )
```

And it works perfectly in the first try! Nah! This take me some time in the 
[jsfiddle](https://jsfiddle.net/gh/get/library/pure/highcharts/highcharts/tree/master/samples/highcharts/demo/dynamic-update).

> All according to keikaku.

# Improving it

We can improve the chart doing for example:

- Make a `plotLine` in y axis at the value 0 to show as reference.

- Make a second yAxis to the right showing the last generated value, this 
can be done using the next function in the `tickPositioner`argument.

```{r}
tick_post_fn <- "function(min,max){        
  var data = this.chart.yAxis[0].series[0].processedYData;
  //last point
  return [Math.round(1000 * data[data.length-1])/1000]; 
}"
```

- Enable the navigator panel `hc_navigator(enabled = TRUE)`.

- We can put an informative title showing the specification of the simulated
model.

```{r}
formula <- purrr::map2(ar, seq_along(ar), function(par, t){
  
  if(par > 0 & t > 1) {
    par <- str_c("+ ", par)
  }
  
  stringr::str_glue("{ par } \\times X_{{ t - { i } }}", i = t)
  
  }) 
  
formula <- purrr::reduce(formula, str_c, sep = " ")

formula <- str_c("$$ X_t = ", formula, " + \\epsilon_t$$")
```

```{r, results='asis'}
cat(formula)
```

Alternatively, formula using subscript and superscript tags.

```{r}
formula <- purrr::map2(ar, seq_along(ar), function(par, t){
  
  if(par > 0 & t > 1) {
    par <- str_c("+ ", par)
  }
  
  htmltools::tagList(par, "×", "X", tags$sub(stringr::str_glue("t - { i }", i = t)))
  
}) 

formula <- purrr::reduce(formula, htmltools::tagList)

formula <- str_c("X", tags$sub("t") %>% as.character(), " = ",formula %>% as.character(), " + &epsilon;",  tags$sub("t") %>% as.character())
```

```{r, results='asis'}
cat(formula)
```

- Helper button to remove last values.

```{r}
rm_poinst_fn <- "function () {
  for (var i = 1; i <= 500; i++) {
    if (this.series[0].data.length) {
      this.series[0].data[0].remove();
    }
  }
}"
```

- And others tweaks.

```{r}
highchart() %>% 
  hc_add_series(data = ts, name = "Process") %>% 
  hc_chart(
    events = list(load = JS(load_fn_glued)), 
    animation = list(duration = time*1000/2)
    ) %>% 
  hc_title(text = "Autoregressive process") %>%
  hc_subtitle(text = as.character(formula), useHTML = TRUE) %>% 
  hc_plotOptions(series = list(marker = list(enabled = FALSE))) %>% 
  hc_tooltip(valueDecimals = 3) %>% 
  hc_xAxis(width = "95%") %>% 
  hc_exporting(
    enabled = TRUE,
    buttons = list(
      list(
        text =  "Remove last 500 values",
        onclick = JS(rm_poinst_fn),
        theme = list(stroke = 'silver')
        )
      )
    ) %>% 
  hc_yAxis_multiples(
    # default axis
    list(
      title = list(text = ""),
      plotLines = list(
        list(value = 0, width = 2, color = "#AAA", zIndex = 1)
        )
      ),
    # opposite axis
    list(
      title = list(text = ""),
      linkedTo = 0,
      opposite = TRUE,
      gridLineWidth = 0,
      tickPositioner = JS(tick_post_fn)
    )
  ) %>% 
  hc_navigator(
    enabled = TRUE, 
    series = list(type = "line"),
    xAxis = list(labels = list(enabled = FALSE), width = "95%")
  ) 
```

# Extending it

We can encapsulate all this code into a function:

```{r}
library(highcharter)
library(stringr)

sim_ar_hc <- function(ar = c(0.3, 0.2), time = 1){
  
   # the first values of the model
  ts <- as.vector(arima.sim(model = list(ar = ar), n = 10))
  ts <- round(ts, 3)
  ts
  
load_fn <- "function () {{

	dot = (a, b) => a.map((x, i) => a[i] * b[i]).reduce((m, n) => m + n);
  function randn_bm() {{
  	var u = 0, v = 0;
    while(u === 0) u = Math.random(); //Converting [0,1) to (0,1)
    while(v === 0) v = Math.random();
    return Math.sqrt( -2.0 * Math.log( u ) ) * Math.cos( 2.0 * Math.PI * v );
   }}
  
	var dat = [{ data }];
  var ar = [{ ar }];
  dat = dat.slice(-ar.length);
  
  // set up the updating of the chart each second
  var series = this.series[0];
  
  setInterval(function () {{
      
      console.log(dat);
      var new_value = dot(dat, ar) + randn_bm();
      new_value = Math.round(1000 * new_value)/1000
      console.log(new_value)
      dat.shift(); 
      dat.push(new_value);
      series.addPoint([new_value]);  
      
      //if (series.data.length < 500) {{
      //  series.addPoint([new_value], true, false);
      //}} else {{
      //  series.addPoint([new_value], true, true);
      //}}
      
  }}, { time });
  
}}"
  
  load_fn_glued <- str_glue(
    load_fn,
    data = str_c(ts, collapse = ","),
    ar = str_c(ar, collapse = ","),
    time = time * 1000
  )
  
  tick_post_fn <- "function(min,max){        
    var data = this.chart.yAxis[0].series[0].processedYData;
    //last point
    return [Math.round(1000 * data[data.length-1])/1000]; 
  }"
  
  rm_poinst_fn <- "function () {
    for (var i = 1; i <= 500; i++) {
      if (this.series[0].data.length) {
        this.series[0].data[0].remove();
      }
    }
  }"
    
  formula <- purrr::map2(ar, seq_along(ar), function(par, t){
    
    if(par > 0 & t > 1) {
      par <- str_c("+ ", par)
    }
    
    htmltools::tagList(par, "×", "X", tags$sub(stringr::str_glue("t - { i }", i = t)))
    
  }) 
  
  formula <- purrr::reduce(formula, htmltools::tagList)
  
  formula <- str_c("X", tags$sub("t") %>% as.character(), " = ",formula %>% as.character(), " + &epsilon;",  tags$sub("t") %>% as.character())
  
  hc <- highchart() %>% 
    hc_add_series(data = ts, name = "Process") %>% 
    hc_chart(
      events = list(load = JS(load_fn_glued)), 
      animation = list(duration = time*1000/2)
      ) %>%
    hc_title(text = "Autoregressive process") %>%
    hc_subtitle(text = formula, useHTML = TRUE) %>%
    hc_plotOptions(series = list(marker = list(enabled = FALSE))) %>%
    hc_tooltip(valueDecimals = 3) %>%
    hc_xAxis(width = "95%") %>%
    hc_exporting(
      enabled = TRUE,
      buttons = list(
        list(
          text =  "Remove last 500 values",
          onclick = JS(rm_poinst_fn),
          theme = list(stroke = 'silver')
          )
        )
      ) %>%
    hc_yAxis_multiples(
      # default axis
      list(
        title = list(text = ""),
        plotLines = list(
          list(value = 0, width = 2, color = "#AAA", zIndex = 1)
          )
        ),
      # opposite axis
      list(
        title = list(text = ""),
        linkedTo = 0,
        opposite = TRUE,
        gridLineWidth = 0,
        tickPositioner = JS(tick_post_fn)
      )
    ) %>%
    hc_navigator(
      enabled = TRUE,
      series = list(type = "line"),
      xAxis = list(labels = list(enabled = FALSE), width = "95%")
    )
  
  hc
  
}
```

White Noise:

```{r}
sim_ar_hc(ar = c(0))
```


Some traditional AR model:

```{r}
sim_ar_hc(ar = c(.8, .1))
```

# Let it shine (with {shiny} package)

In this case we can update the highcharts using the new set of proxy functions. 
You can check the code of the shiny app in https://github.com/jbkunst/highcharter-shiny/blob/master/05-arma/app.R and the 
app is in https://jbkunst.shinyapps.io/05-arma/?showcase=0

<!-- <iframe src="https://jbkunst.shinyapps.io/05-arma/?showcase=0" class="l-page" height="750px"></iframe> -->

```{r, class.source = "l-page", echo=FALSE}
knitr::include_app("https://jbkunst.shinyapps.io/05-arma/?showcase=0", height = "750px")
```

# References

- https://jsfiddle.net/gh/get/library/pure/highcharts/highcharts/tree/master/samples/highcharts/demo/dynamic-update
- https://jsfiddle.net/BlackLabel/evypfr1L/
- http://jsfiddle.net/upt4cbqj/
- http://www.java2s.com/Tutorials/highcharts/Example/Series_Data/Remove_first_data_item_from_series.htm
