---
title: "Let's party with the new ggparty package"
description: |
  The new package ggparty package is a lifesaver when you want yo show your 
  tree models to others who surely don't like the old base plot looking. And...
  You know, this is a helper post for my future self who have some
  memory problems. 
  The ggparty package have a lot of functions to customize your tree plots much 
  as you want so we'll see the most important (for me and my future self) here.
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 07-22-2019
output:
  distill::distill_article:
    self_contained: true
    toc: true
draft: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source(here::here("_R/blog_setup.R"))
knitr::opts_chunk$set(layout = "l-body")
```


## Introduction

Surely more than once you have needed to make a decision tree because:

- _Someone_ or _they_ asked you.
- Is an easy way to explain how the patterns in the data works, because trees are 
usually the first approach to machine learnig for a lot of us.
- Just you like them/it. Period.

If you have some experience with decision trees in R you know the most popular packages 
are `partykit` and `rpart`. In personal experience I always have much better results in `partykit` compared with `rpart` in terms of splits/structure. Besides the default plot is better IMHO, but that doesn't mean is a good looking plot _hahaha_:

```{r}
library(partykit)

irisct <- ctree(Species ~ .,data = iris)

plot(irisct, gp = gpar(fontsize = 8))
```

That is 80's flavor :D.

## Data, duh!

I pass my days in a chilean bank so usually work with data similar to the `germarcredit.csv`:
a set of variables and a reponse varialbe indicating if the given account was in default or not.

```{r}
library(tidyverse)

credit <- read_csv(here::here("data/credit.csv"))

set.seed(123)

credit <- credit %>% 
  group_by(bad) %>% 
  sample_n(10000, replace = TRUE) %>% 
  ungroup() %>% 
  select(-id_client, -contains("phone"),  -contains("residence"),
         -profession_code, -payment_day, -marital_status, -sex,) %>% 
  mutate(bad = ifelse(bad == 1, "bad", "good")) %>% 
  mutate_if(is.character, as.factor)
 

glimpse(credit)
```

## Back to the party

Let's start with a simple model using the previous data and then plot it:

```{r}
tree <- ctree(bad ~ ., data = credit, control = ctree_control(maxdepth = 3))

plot(tree, gp = gpar(fontsize = 8))
```

Now the `ggparty`'s version:

```{r}
library(ggparty)

# from the example of ?ggparty
ggparty(tree) +
  geom_edge() +
  geom_edge_label() +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  geom_node_label(aes(label = info), ids = "terminal")
```

Oh! It looks promising but is kind of far of what `ggparty` can do.  

The firsts questions about how `ggparty` works are:

> Where does **splitvar** or **info** come from? - Me.

> Why the terminal node show (in this case) **NA** - Me too.

And the answers are:

```{r}
the_magic <- as_tibble(ggparty(tree)$data)

glimpse(the_magic)
```

In summary:

- The function `ggparty::ggparty` parse the tree (a `party` object) to a `data.frame` so
you can use ggplot to show the parsed data.
- Each row is a node with the related informaction: the `id`, position (`x` and `y`),
`parent`, `kids`, etc.
- And the magic keep going with the `nodedata_*` colums which have the data related in 
that node!

That's why if you see the `info` column, you can notice is full of `NA`s and then 
the part `label = info` is showing us just the same `NA`s.

So, we can do? A lot! In fact, if you are if you are familiarized with `tibble`s and 
some `purrr` functions you now is easy operate list-columns.


```{r}
library(scales) # for comma,percent function

ggparty(tree) +
  geom_edge() +
  geom_edge_label() +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  geom_node_label(aes(label = info), ids = "terminal") +
  geom_node_label(
    aes(
      label = str_c(
        "N: ",
        comma(nodesize),
        "\nTarget rate: ",
        percent(map_dbl(nodedata_bad, ~ mean(.x == "bad")))
        )
      ),
  size = 2,
  ids = "terminal")
```

Mmm... Too much to do! Right?







