---
title: "Regression, ANOVA, t-test are related..."
description: |
  In a new episode of things I forgot to find.
author:
  - name: Joshua Kunst
    url: http://jkunst.com/
date: 06-08-2021
output:
  distill::distill_article:
    self_contained: true
categories:
  - minipost
  - stats
preview: images/preview.png  
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

_Source_: https://stats.stackexchange.com/questions/59047/how-are-regression-the-t-test-and-the-anova-all-versions-of-the-general-linear

I Always fail in remember the code to show how this models are related, so I
will put here for my future me. The importan thing is check the p-values.

The data, according `help(sleep)`:

> Data which show the effect of two soporific drugs (increase in hours of sleep compared to control) on 10 patients -Scheffé, Henry (1959) The Analysis of Variance. New York, NY: Wiley.

Now, load packages and data.

```{r}
library(tibble)
library(broom)

data("sleep")

sleep <- as_tibble(sleep)

glimpse(sleep)
```

## Regression

```{r}
linear_model <- lm(extra ~ group, data = sleep)

summary(linear_model)

tidy(linear_model)
```

## ANOVA

```{r}
anova <- aov(extra ~ group, data = sleep)

summary(anova)

tidy(anova)
```

## $t$-test

```{r}
t_test <- t.test(extra ~ group, var.equal = TRUE, data = sleep) 

t_test

tidy(t_test)
```